---
title: "CPLN 675 Midterm"
author: "Marquise Williams and Will Friedrichs"
date: "3/29/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    runtime: shiny
---

For this project, a generalized linear model will be used to predict flood inundation zones first in Calgary, then in Sacramento County. For each area, predictions will estimate whether or not a given 500m by 500m "cell" of area can be described as in a flood inundation zone. This prediction process is described in four steps: data gathering, building variables, logistic metrics, and goodness of fit metrics.

```{r setup, include=TRUE,message = FALSE,cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
sf::sf_use_s2(FALSE)
options(scipen=999)
library(caret)
library(pscl)
library(plotROC)
library(pROC)
library(sf)
library(tidyverse)
library(knitr)
library(kableExtra)
library(tigris)
library(viridis)
library(terra)
library(cowplot)
library(modelr)
```

## 1) Goals, & Motivations, & Model Description 
### Goals, & Motivations, & Model Description 

Goals:
This document serves as supplemental material to the proposal for the creation of municipal Flood Inundation Modeling & Planning Centers (FIMPC). The purpose of this document is provide an predictive 'model' that determines the probability an area floods. The aim is to provide information that helps residents, property owners, planners and municipalities make cost-effective decisions regarding their potential loss estimates based on the severity of flood active in their area.

Motivations: 
The importance is that this ‘model’ can be used by several local governments to assess their communities for flood risk and response with efforts to proactively prepare against potential flood damage based upon the likelihoods of outcomes. Specially, this ‘model’ has the potential to save lives, money and resources on the preventative side for jurisdictions that are alike.

Model Description:
The data collected for ‘models’ would include current movement or flows of streams, and the types of land use surfaces that exist in communities such as residential, commercial, agricultural, because they often determine the likelihood that the land surfaces would be water permeable. Also, data on community features such as the municipal boundaries, major city locations, urban tree cover, and distance to major rivers and streams are included to provide a physical, natural, and hydrological context to the predictions. 

The document walks through the generalized linear ‘model’ used to predict flood inundation zones first in Calgary, then in takes the same approach to predicting flooding risk areas for Sacramento County, which would be the goal of the municipal FIMPC. For each area, predictions will estimate whether a given 500m by 500m "cell" of area can be described as in a flood inundation zone or at risk of flooding.

This prediction process is described in four steps: data gathering, building variables, logistic metrics, and goodness of fit metrics.




Goals:
This document serves as supplemental material to the proposal for the creation of municipal Flood Inundation Modeling & Planning Centers (FIMPC). The purpose of this document is provide an predictive 'model' that determines the probability an area floods. The aim is to provide information that helps residents, property owners, planners and municipalities make cost-effective decisions regarding their potential loss estimates based on the severity of flood active in their area.

Motivations: 
The importance is that this ‘model’ can be used by several local governments to assess their communities for flood risk and response with efforts to proactively prepare against potential flood damage based upon the likelihoods of outcomes. Specially, this ‘model’ has the potential to save lives, money and resources on the preventative side for jurisdictions that are alike.

Model Description:
The data collected for ‘models’ would include current movement or flows of streams, and the types of land use surfaces that exist in communities such as residential, commercial, agricultural, because they often determine the likelihood that the land surfaces would be water permeable. Also, data on community features such as the municipal boundaries, major city locations, urban tree cover, and distance to major rivers and streams are included to provide a physical, natural, and hydrological context to the predictions. 

The document walks through the generalized linear ‘model’ used to predict flood inundation zones first in Calgary, then in takes the same approach to predicting flooding risk areas for Sacramento County, which would be the goal of the municipal FIMPC. For each area, predictions will estimate whether a given 500m by 500m "cell" of area can be described as in a flood inundation zone or at risk of flooding.

This prediction process is described in four steps: data gathering, building variables, logistic metrics, and goodness of fit metrics.




## 2) Building Variables and Running Model

### Variable descriptions

The following factors will be examined as predictors of Calgary's flood inundation zones.

<code>cal_dem</code>, <code>sac_dem</code>: For Calgary and Sacramento County respectively, these represent elevation data relative to the lowest point in the area. Flood inundation is probably more likely in low-lying areas than high elevation areas, so there is reason to suspect this factor might be significant.

<code>cal_streamdist</code>, <code>sac_streamdist</code>: For Calgary and Sacramento County respectively, these describe the distance between any cell and a stream. 

<code>cal_riverdist</code> and <code>sac_riverdist</code>: For Calgary and Sacramento County respectively, this describes the distance the cells are to major rivers. These variables come to mind considering our hypothesis that areas with streams have a higher flooding susceptibility. 

Additionally, the the type of land cover may also factor into whether an area is susceptible to flooding. For example certain land uses, such as commercial or industrial may allow for greater soil impermeability than others, which increases vulnerability to flooding in some urban contexts without proper drainage. 

For comparative reasons, land uses have been condensed into five land use zones categories for Calgary & Sacramento County: 

<code>cal_residcover</code> and <code>sac_residcover</code>: This describes areas of Calgary and Sacramento County with residential land uses.

<code>cal_mixedcover</code> and <code>sac_mixedcover</code>: This describes areas of Calgary and Sacramento County with mixed Use or transit orientated development  land uses.

<code>cal_commecover</code> and <code>sac_commecover</code>: This describes areas of Calgary and Sacramento County with commercial land uses.

<code>cal_induscover</code> and <code>sac_induscover</code>: This describes areas of Calgary and Sacramento County with industrial land uses.

<code>sac_agricover</code> and <code>sac_agricover</code>: This describes areas of Calgary and Sacramento County with agricultural & open space land uses. 

<code>cal_treecover</code> and <code>sac_treecover</code>: These are used to describe the tree cover within Calgary and Sacramento County respectively. The presence of tree density may play a factor in water absorption as well as soil permeability.

<code>cal_inundation</code>: This is what our model aims to predict. We will train our model on part of Calgary flood inundation fishnet, and test our predictions on another part of it.

```{r load_json, warning = FALSE, message = FALSE}
cal_dem <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/Cal_dem_filled.json", quiet = TRUE)

cal_streamdist <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/cal_dist2streams.json", quiet = TRUE)

cal_inundation <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/calgary_inundation.json", quiet = TRUE)

cal_inundation[is.na(cal_inundation)] = 0

sac_dem <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/sac_dem_filled.geojson", quiet = TRUE)

sac_streamdist <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/sac_dist2streams.geojson", quiet = TRUE)

##Sacramento & Calgary Distance to Major Rivers Variables

sac_riverdist <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacramentoRiverDistance.json", quiet = TRUE) 

cal_riverdist <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryRiverDistance.json", quiet = TRUE) 

##Sacramento & Calgary Land Use Variables

sac_residcover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_Resi_UPDATED.json", quiet = TRUE) ##Residential land Use

sac_mixedcover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_mixed_filled.geojson", quiet = TRUE) ## Mixed Use/ Transit Orientated Development Land Use

sac_commecover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_mixed_filled.geojson", quiet = TRUE) ## Commercial Land Use

sac_induscover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_industrial_filled.geojson", quiet = TRUE) ##Industrial Land Use

sac_agricover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_argric_filled.geojson", quiet = TRUE) ## Agricultural & Open Space Land Use

## Calgary Land Use
cal_residcover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_resid.json", quiet = TRUE)  ##Residential land Use

cal_mixedcover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_mixed.json", quiet = TRUE) ## Mixed Use/ Transit Orientated Development Land Use

cal_commecover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_commer.json", quiet = TRUE) ## Commercial Land Use

cal_induscover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_industr.json", quiet = TRUE) ##Industrial Land Use

cal_agricover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_agric.json", quiet = TRUE) ## Agricultural & Open Space Land Use


## Tree Cover
cal_treecover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/CalgaryFishnet_TreeCover__.json", quiet = TRUE)

sac_treecover <- st_read("https://raw.githubusercontent.com/Willfriedrichs/Midterm_Repo/main/Finalized_Layers/SacFN_New_TreeCover.json", quiet = TRUE)

```

### Putting variables together

The Calgary variables are joined together to form one table that includes each of the independent variables and the dependent variable. These variables will also be adjusted for better comparison across regions. For example, some variables only need their geometry and the metric that is measures such as mean distance to stream from any given cell or majority of land use type for cells.

```{r load_chesco, warning = FALSE, message = FALSE}
##Calgary Model

# cal_dem only needs its geometry and information as to the mean elevation of each cell.
# mean elevation will be re-evaluated relative to minimum non-zero fishnet cell
cal_dem = cal_dem %>% select(ZonalSt_Fishnet1.MEAN)
min_nonzero_cell = min(cal_dem$ZonalSt_Fishnet1.MEAN)
cal_dem$ZonalSt_Fishnet1.MEAN = cal_dem$ZonalSt_Fishnet1.MEAN - min_nonzero_cell

#cal_streamdist only needs its geometry and the mean distance to stream from any given cell.
cal_streamdist = cal_streamdist %>% select(dist2streams.MEAN)

#cal_inundation only needs its geometry and the majority inundation for cells
cal_inundation = cal_inundation %>% select(calgary_inundation.MAJORITY) 

#cal_residcover only needs its geometry and the majority of the areas pertaining to residential land use for cells
cal_residcover = cal_residcover %>% select(ZonalSt_Calgary1_MAJORITY) %>% st_set_crs(4326) %>% st_transform(4326)

#cal_mixedcover only needs its geometry and the majority of the areas pertaining to mixed Use or transit orientated development land use for cells
cal_mixedcover = cal_mixedcover %>% select(ZonalSt_Calgary2_MAJORITY) %>% st_set_crs(4326) %>% st_transform(4326)

#cal_commecover only needs its geometry and the majority of the areas pertaining to  commercial land uses for cells 
cal_commecover = cal_commecover %>% select(ZonalSt_Calgary3_MAJORITY) %>% st_set_crs(4326) %>% st_transform(4326)

#cal_induscover only needs its geometry and the majority of the areas pertaining to industrial land uses for cells
cal_induscover = cal_induscover %>% select(ZonalSt_Calgary4_MAJORITY) %>% st_set_crs(4326) %>% st_transform(4326)

#cal_agricover only needs its geometry and the majority of the areas pertaining to  agricultural & open space land uses for cells
cal_agricover = cal_agricover %>% select(ZonalSt_Calgary5_MAJORITY) %>% st_set_crs(4326) %>% st_transform(4326) 

#cal_treecover only needs its geometry and the count of trees in the area for each cell 
cal_treecover = cal_treecover %>% select(ZonalSt_Calgary6_COUNT) %>% st_set_crs(4326) %>% st_transform(4326)
cal_treecover[is.na(cal_treecover)] <- 0

#cal_riverdist only needs its geometry and the mean distance to major rivers from any given cell 
cal_riverdist = cal_riverdist %>% select(LastTableCalgary.MEAN) %>% st_set_crs(4326) %>% st_transform(4326)

##Sacramento Model

# sac_dem only needs its geometry and information as to the mean elevation of each cell.
# since the minimum mean elevation is nonzero in Sacramento, no readjustment is necessary.
sac_dem = sac_dem %>% select(demFilledSAC.MEAN)

#sac_streamdist only needs its geometry and the mean distance to stream from any given cell.
sac_streamdist = sac_streamdist %>% select(dist2streamsSAC.MEAN)

#sac_residcover only needs its geometry and the majority of the areas pertaining to residential land use for cells
sac_residcover = sac_residcover %>% select(ZonalSt_SacFN_N7_MAJORITY)

#sac_mixedcover only needs its geometry and the majority of the areas pertaining to mixed Use or transit orientated development land use for cells
sac_mixedcover = sac_mixedcover %>% select(ZonalSt_SacFN_N2_MAJORITY)

#sac_commecover only needs its geometry and the majority of the areas pertaining to  commercial land uses for cells
sac_commecover = sac_commecover %>% select(ZonalSt_SacFN_N2_MAJORITY)

#sac_induscover only needs its geometry and the majority of the areas pertaining to industrial land uses for cells
sac_induscover = sac_induscover %>% select(ZonalSt_SacFN_N4_MAJORITY)

#sac_agricover only needs its geometry and the majority of the areas pertaining to  agricultural & open space land uses for cells
sac_agricover = sac_agricover %>% select(ZonalSt_SacFN_N5.MAJORITY)

#sac_treecover only needs its geometry and the count of trees in the area for each cell 
sac_treecover = sac_treecover %>% select(ZonalSt_SacFN_N5_COUNT)

#sac_riverdist only needs its geometry and the mean distance to major rivers from any given cell 
sac_riverdist = sac_riverdist %>% select(ZonalSt_SacFN_N1.MEAN)

# Once in a single dataframe, discard OIDs and rename variables
##All Calgary

allCalgary = st_join(cal_dem, cal_streamdist, largest = TRUE) %>%
             st_join(cal_inundation, largest = TRUE) %>%
             st_join(cal_residcover, largest = TRUE) %>%
             st_join(cal_treecover, largest = TRUE) %>%
             st_join(cal_commecover, largest = TRUE) %>%
             st_join(cal_induscover, largest = TRUE)


allCalgary = allCalgary %>% rename(meanElevation = ZonalSt_Fishnet1.MEAN,
                                   meanStreamDistance = dist2streams.MEAN,
                                   inundation = calgary_inundation.MAJORITY,
                                   redisCover = ZonalSt_Calgary1_MAJORITY,
                                   commecover = ZonalSt_Calgary3_MAJORITY,
                                   induscover = ZonalSt_Calgary4_MAJORITY,
                                   treecover = ZonalSt_Calgary6_COUNT)


```


Each of the predictors is mapped below.

```{r ggplots, warning = FALSE, message = FALSE}

ggplot() +
  geom_sf(data=allCalgary, aes(fill=inundation, color=inundation)) +
  labs(title="Inundation") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )
ggplot() +
  geom_sf(data=allCalgary, aes(fill=as.numeric(meanElevation), 
                              color=as.numeric(meanElevation))) +
  labs(title="Elevation") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )

ggplot() +
  geom_sf(data=allCalgary, aes(fill=as.numeric(meanStreamDistance),                       
                               color=as.numeric(meanStreamDistance))) +
  labs(title="Distance to Stream") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )


ggplot() +
  geom_sf(data=allCalgary, aes(fill=redisCover,
                               color=redisCover)) +
  labs(title=":LU Residential") +
  theme(axis.text.x=element_blank(), #remove x axis labels
    axis.ticks.x=element_blank(), #remove x axis ticks
    axis.text.y=element_blank(),  #remove y axis labels
    axis.ticks.y=element_blank()  #remove y axis ticks
    )


ggplot() +
  geom_sf(data=allCalgary, aes(fill=commecover,
                               color=commecover)) +
  labs(title="LU Commercial") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )


ggplot() +
  geom_sf(data=allCalgary, aes(fill=induscover,
                               color=induscover)) +
  labs(title="LU Industrial") +
  theme(axis.text.x=element_blank(), #remove x axis labels
    axis.ticks.x=element_blank(), #remove x axis ticks
    axis.text.y=element_blank(),  #remove y axis labels
    axis.ticks.y=element_blank()  #remove y axis ticks
    )

  

```

The variables are now ready to be split into training and test sets, and finally, added to the linear model. The <code>set.seed()</code> and <code>createDataPartition</code> functions help to randomly select a portion (in this case 70% of cells) to join the training set. The rest form the test set.

```{r training_set}
# Randomly separate data into 70% training set, 30% test set
set.seed(3456)
trainIndex <- createDataPartition(allCalgary$inundation, p = .70,
                                  list = FALSE,
                                  times = 1)

allCalgaryTrain <- allCalgary[ trainIndex,]
allCalgaryTest  <- allCalgary[-trainIndex,]
```
                     
The <code>glm()</code> takes in the independent variables and dependent from a portion of the dataset and produces a formula that minimizes prediction error. The formula's performance predicting the dependent variables in the test set can be used to characterize the performance of the model in Calgary. It can also be used to predict inundation given the previously outlined independent variables in other locations, such as Sacramento.

```{r create_model, warining = FALSE, message = FALSE}
# Convert to regular dataframe, R cannot run a regression on an sf
inundationModel <- glm(inundation ~ ., 
                    family="binomial"(link="logit"), data = allCalgaryTrain %>%
                                                            as.data.frame() %>%
                                                            select(-geometry))
```


## 3) Model results, goodness of fit metrics, and cross validation

### Model results

The summary below displays information describing the relationship between the independent variables and inundation zones in Calgary. Note that not all of the data collected is included in the final model. Some factors like mixed use and agricultural land cover were not at all significant predictors of inundation zones, and others, such as distance to rivers, were left out due to multicolinearity.

```{r logistc_metrics, warining = FALSE, message = FALSE}
# Get summary statistics for the inundationModel
summary(inundationModel)
```

The model output, as displayed by the <code>summary()</code> function, shows the coefficients of the model. This is mathematically what produces inundation predictions by cell. The coefficients of <code>meanElevation</code> and <code>meanStreamDistance</code> are both negative, meaning that the greater these values get, the lower the chance a given cell will be predicted to be in an inundation zone. This makes sense, as low elevation and low distance to streams pose greater flood risks. <code>redisCover</code>, <code>commecover</code>, and <code>induscover</code> are all also negatively related to the output, meaning the model estimates these types of development to decrease the likelihood of a cell being in an inundation zone.

The <code>inundationModel</code> can be used to predict for any geographic cells given the previously specified independent variables with the <code>predict()</code> function. In this case, the model will be used to predict for the test set. The output, <code>inundationProbability</code>, will describe the model's prediction of the likelihood that each cell will be in an inundation zone.

The histogram below demonstrates that the great majority of tracts are predicted to have a low likelihood of being in an inundation zone. This makes sense, because within Calgary city limits, based on the training set, most cells of land were not in an inundation zone.

```{r predict_calgary_test}
# Get the results of the model on the Test set, create a histogram of the probability
# The parameter type="response" ensures that the probabilities range from 0 to 1.
inundationTrainProb <- predict(inundationModel, allCalgaryTest, type="response")

hist(inundationTrainProb)
```

The two charts below display inundation predictions for the test set. The first chart shows predictions for cells that ended up being non-inundation (0) zones, and the second chart shows predictions for inundation (1) zone cells. The y-axis describes quantities of predictions produced for each level of probability.

The model generally predicts much lower probabilities in actual non-inundation zones. In actual inundation zones, inundation zone likelihood estimates are more likely to be higher. They do not pass 0.5, though, signaling the model's estimation that even given its inputs, it never is more than 50% sure that a cell is in an inundation zone.

```{r plot_preds}
inundationTestProb <- data.frame(obs = as.numeric(allCalgaryTest$inundation),
                        pred = inundationTrainProb)

ggplot(inundationTestProb, aes(x = pred, fill=as.factor(obs))) + 
  geom_density() +
  facet_grid(obs ~ .) + 
  xlab("Probability") + 
  geom_vline(xintercept = .5) +
  scale_fill_manual(values = c("dark green", "dark blue"),
                      labels = c("No inundation zone predicted","Inundation zone predicted"),
                      name = "") + 
  labs(title="Test set: predicted probabilities by actual inundation zone status")
```


### Goodness of fit metrics

Goodness of fit metrics quantify a models predictive performance. A confusion matrix takes predictions and true values and delivers the count of true negatives, false negatives, true positives, and false positives. 

In order to create a confusion matrix, a threshold for predicting the dependent variable must be selected. A threshold of 50% means that when the model delivers a prediction of 0.5 or more, it becomes a prediction of an inundation zone.

A threshold of 0.15 produces the confusion matrix below. Due to the low likelihood of inundation predicted by the model for all zones, a low threshold is necessary to achieve any sensitivity (true positivity). This means some specificity is sacrificed, as more non-inundation zones are labeled as inundation zones.

```{r confusion_matrix, message = FALSE, warning = FALSE}
inundationTestProb$predClass  = ifelse(inundationTestProb$pred > .15 ,1,0)

caret::confusionMatrix(reference = as.factor(inundationTestProb$obs), 
                       data = as.factor(inundationTestProb$predClass), 
                       positive = "1")
```

An ROC or "Reciever Operating Characteristic" curve is anther metric that describes a model's goodness of fit. It measures the relationship between sensitivity (True positive fraction) and specificity (false positive fraction) at different levels of prediction threshold. On the chart below, the black line is the performance of the model, and it is compared to a gray line that represents a hypothetical "coin flip" model.

```{r roc_curve, message = FALSE, warning = FALSE}

ggplot(inundationTestProb, aes(d = obs, m = pred)) + 
  geom_roc(n.cuts = 50, labels = FALSE) + 
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = 'grey') 
```

The AUC (Area Under Curve) measures the models ability to make distinctions between the two classes, in this case non-inundation and inundation zones on a fishnet. An AUC close to 1.0 suggests that the model is a relatively well-performing one.
  
```{r auc, warning = FALSE}
auc(inundationTestProb$obs, inundationTestProb$pred)
```

### Cross validation

Cross validation tests the model on different combinations of data other than just a single test set. In this instance of 100-fold validation, the model will run across 100 different divisions of test and training sets. Summary statistics of the accuracy of each of the 100 sets of predictions are displayed below. The results show a consistent level of accuracy, with a mean slightly above 0.95 out of 1.0.

```{r k_fold, warning = FALSE, message = FALSE}
ctrl <- caret::trainControl(method = "cv", 
                     number = 100, 
                     savePredictions = TRUE)

cvFit <- caret::train(as.factor(inundation) ~ .,  
                 data = allCalgaryTrain %>% 
                 as.data.frame() %>%
                 select(-geometry), 
               method="glm",
               family="binomial",
               trControl = ctrl)

summary(cvFit$resample$Accuracy)

```

### Spatial cross validation

Spatial cross validation creates regional test sets, such that any region of the fishnet where the fuction performs particularly poorly can be identified. Summary statistics of the accuracy of each of the 100 spatially-specific predictions are displayed below. These results, similar to those above, confirm the model 

```{r spatial_CV, warning = FALSE, message = FALSE}
library(mlr)

# https://www.r-bloggers.com/2018/07/visualization-of-spatial-cross-validation-partitioning/
# https://geocompr.robinlovelace.net/spatial-cv.html

# create a resampling description with parameters on how we will resample our data for the cross validation. 

factorized_training = st_drop_geometry(allCalgaryTrain)
factorized_training$treecover[is.na(factorized_training$treecover)] <- 0
factorized_training$inundation = as.factor(factorized_training$inundation)

calgaryTrainingTask = makeClassifTask(data = factorized_training,
                                      target = "inundation",
                                      coordinates = as.data.frame(st_coordinates(st_centroid(allCalgaryTrain)))
)

lrn = makeLearner(cl = "classif.binomial",
                  link = "logit",
                  predict.type = "prob",
                  fix.factors.prediction = TRUE)

perf_level = makeResampleDesc(method = "SpRepCV", folds = 5, reps = 100)

set.seed(012348)
sp_cv = mlr::resample(learner = lrn, task = calgaryTrainingTask,
                      resampling = perf_level, 
                      measures = mlr::acc)

summary(sp_cv$measures.test$acc)
```

## 4) Confusion matrix map, Calgary inundation predictions, and predictions for Sacramento

### Confusion matrix map

The maps below show confusion matrix values by cell across Calgary based on the model.

```{r confusion_mapped, warning = FALSE, message = FALSE}
# Combine the Calgary training and test sets, including associated predictions
inundationAllProb <- add_predictions(allCalgary, inundationModel, type="response")

inundationAllProb$confusion = "true negative"
inundationAllProb <- transform(inundationAllProb, confusion = ifelse(pred >= 0.15 & 
                                                                     inundation == 1, 
                                                                     "true positive", 
                                                                     confusion))
inundationAllProb <- transform(inundationAllProb, confusion = ifelse(pred >= 0.15 & 
                                                                     inundation == 0, 
                                                                     "false positive", 
                                                                     confusion))
inundationAllProb <- transform(inundationAllProb, confusion = ifelse(pred < 0.15 & 
                                                                     inundation == 1, 
                                                                     "false negative", 
                                                                     confusion))

ggplot() +
  geom_sf(data=inundationAllProb, aes(fill=confusion, color=confusion)) +
  labs(title="Confusion Matrix by Cell") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )
```

```{r full_cal_prediciton, warning = FALSE, message = FALSE}
ggplot() +
  geom_sf(data=inundationAllProb, aes(fill=pred, color=pred)) +
  labs(title="Confusion Matrix by Cell") +
  theme(axis.text.x=element_blank(), #remove x axis labels
      axis.ticks.x=element_blank(), #remove x axis ticks
      axis.text.y=element_blank(),  #remove y axis labels
      axis.ticks.y=element_blank()  #remove y axis ticks
      )

```

